In-silico molecular enrichment and clearance of the human intracranial space

### 2024-12-20 paper/supplementary_main.tex
Remove obsolete comment.
Update Table 2 and put it back in with Methods.

\subsection*{Overview of in-silico scenarios and experiments considered}

See~\Cref{tab:scenarios} for an overview of the different in-silico
scenarios and computational models considered.
\begin{table}
  \begin{center}
  \begin{tabular}{ll|llllll|ll}
    \toprule
    & Description & $R_1, R_2$ & $\xi_{\rm PVS-SAS}$ &  $\beta_{\rm efflux}$ & $D$ & $\hat{u}$ & $\mathbf{u}_{SAS}$ & $c_0$ \\
    \midrule
    A & Baseline & $R_2 = 2 R_1$ & $\xi$ & $10^{-4}$ mm$^2$/s\cite{hornkjol2022csf} & $\phi R D^{\rm gad}_{\rm eff}$\cite{sykova2008diffusion, valnes2020apparent} & $\hat{u}_{\rm prod} + \hat{u}_{\rm cardiac}$ & $\mathbf{u}_{\rm prod}$ & 0 \\ 
    B1 & Sheath permeability & $R_2 = 2 R_1$ & $0, \xi, 10 \xi, 100 \xi, 1000 \xi$ & $10^{-4}$ mm$^2$/s\cite{hornkjol2022csf} & $\phi R D^{\rm gad}_{\rm eff}$\cite{sykova2008diffusion, valnes2020apparent} & $\hat{u}_{\rm prod} + \hat{u}_{\rm cardiac}$ & $\mathbf{u}_{\rm prod}$ & 0 \\ 
    B2 & Health permeability & $R_2 = 2 R_1$ & $0, \xi, 10 \xi, 100 \xi, 1000 \xi$ & $10^{-4}$ mm$^2$/s\cite{hornkjol2022csf} & $\phi R D^{\rm gad}_{\rm eff}$\cite{sykova2008diffusion, valnes2020apparent} & $\hat{u}_{\rm prod} +  \hat{u}_{\rm cardiac} + \hat{u}_{\rm vaso}$ & $\mathbf{u}_{\rm prod}$ & 0 \\ 
    %C & Larger PVS & $R_2 = N R_1$ & -- & -- & --  &  --  & -- &  -- \\
    C & Higher PVS flow & -- & $\xi$ & -- & --  & $\hat{u}_{\rm prod} +  \hat{u}_{\rm cardiac} + \hat{u}_{\rm vaso}$  & -- & --  \\
    AA & PVS size & $R_2 = x \beta R_1$ & $\xi$ & -- & -- & -- &  -- & 0 \\ 
    %C & Larger PVS & $R_2 = N R_1$ & -- & -- & --  &  --  & -- &  -- \\
    B & PVS flow & -- & $\xi$ & -- & --  & $\hat{u}_{\rm prod} + \hat{u}_{\rm vaso}$  & -- & --  \\
    C & Clearance & -- & -- & -- & $\alpha D^{\rm gad, \tau, A\beta}$  & $\hat{u}_{\rm prod}$ & -- & 1 \\
    %E & PVS II & $R_2 = 2 R_1$ & $\xi$, -- & -- & --  &  $\hat{u} = \hat{u}_{\rm prod} + \uparrow\uparrow$  & -- & --  \\
    %F & PVS III & $R_2 = 2 R_1$ & $\xi$, -- & -- & $10 D_{\rm PVS}$ &  $\hat{u} = \hat{u}_{\rm prod}$  & -- & --  \\
    %G & Glymphatics & $R_2 = 2 R_1$ & $\xi$, -- & -- & \cite{sykova2008diffusion, valnes2020apparent} &  $\hat{u} = \hat{u}_{\rm prod} + \uparrow$  & $\mathbf{u}_{\rm PAR}$ > 0 & --  \\
    \bottomrule
    \end{tabular}
    \end{center}
  \caption{Overview of models. (-- denotes as immediately above). Each
    PVS cross-section is modelled as a concentric annulus with inner
    radius $R_1$ and outer radius $R_2$ and such that $R_2 = \beta
    R_1$ with $\beta = 2.0$ based on~\cite{eide2024functional}. Let
    $\xi = \dots$ denote the permeability estimated by Koch et
    al.~\cite{koch2023estimates}. Model A and its variations aims to
    quantify the effect of PVS sheaths by varying the permeability
    $\xi$ between the PVS and its surroundings. We consider a fixed
    permeability $\xi_{\rm PVS-PAR}$ based
    on~\cite{koch2023estimates}. \mer{The permeability between the PVS
      and the SAS $\xi_{\rm PVS-SAS}$ must likely be estimated by a
      bit of trial-error from the observations of timings of PVS-SAS
      and SAS from~\cite{eide2024functional}.} \discuss{The
      extracranial solute efflux permeability $\beta_{\rm efflux}$ is
      uniformly distributed over the outer SAS boundary with a
      reasonable rate as baseline\cite{hornkjol2022csf,
        eide2021clinical, ringstad2024glymphatic}.} At baseline, the
    effective diffusion coefficients $D^{\rm x}_{\rm SAS} = D^{\rm
      x}_{\rm PVS}$ of the SAS and PVS represent the free diffusion
    coefficient of Gadubutrol (${\rm x} = {\rm gad}$) in CSF (water),
    and $D^{\rm x}_{\rm PAR}$ the effective diffusion coefficient of
    human cortical tissue, all at body temperature; weighted by a
    dispersion factor $\alpha$ (see Methods). At baseline, we consider
    net flow due to CSF-production in the SAS and PVS:
    $\mathbf{u}_{\rm SAS} > 0$ and $\hat{u} > 0$ (see Methods) but no
    bulk flow in the parenchyma $\mathbf{u}_{\rm PAR} = 0$. Due to the
    timescale for human intracranial tracer transport (minutes to
    hours) compared to typical (cardiac, respiratory, slow vasomotion)
    human intracranial pulsatility (0.1 -- 1 Hz), we consider net
    (constant-in-time) flow fields only in subsequent model variations
    (but see dispersion factor above).  Model B represents
    perivascular pathway scenarios with more rapid flow in the PVS
    induced by vasomotion (PVS I)~\cite{gjerde2023directional}.
    Models C represents a version of the aforementioned scenarios but
    for studying metabolite clearance (instead of solute influx).}
\label{tab:scenarios}
\end{table}

@@

This theoretical formalism~\cite{gjerde2023directional} is defined
relative to a network in the form of a directed, bifurcating tree with
a single supply node/root $i_0$. To extend to a network of cerebral
arteries with multiple supply nodes (such as the basilar and two
internal carotid arteries in the current data
set~\cite{hodneland2019new}), we separate the network $\Lambda$ into
edge-disjoint subnetworks $\Lambda^1, \Lambda^2, \Lambda^3$, one for
each of the supply nodes (\Cref{fig:results1}A, C,
\Cref{fig:app:peristalsis}). Each node is assigned to the subnetwork
associated with the nearest supply node, and edges between nodes are
preserved. Next, we compute a minimal, bifurcating and directed tree
representation of each subnetwork: $\mathcal{T}^1, \mathcal{T}^2,
\mathcal{T}^3$. Each tree $\mathcal{T}^j = \cup E_n^j$ consists of the
subset of the nodes from $\Lambda^j$ that have degree 1 (are leaf or
root nodes) or degree $3$ (are true bifurcation points), and each path
between nodes with degree $\not = 2$ in $\Lambda^j$ is represented in
$\mathcal{T}^j$ by an edge $E^j$ with edge length $L$ corresponding to
the total length of the original path and edge radius $R^j$ as the
average of the path radii. \discuss{For the few (2) nodes in $\Lambda$ with
degree $>3$; i.e.~junctions where a branch splits into more than two
downstream branches, only the two longest downstream paths are
included in $\mathcal{T}^j$.}

For each subtree $\mathcal{T}^j = \cup_n E_n^j$, we compute the
time-averaged downstream flow rate $\langle Q'_n \rangle$ induced by
the vascular wall motion for each edge $n$ via~\cite[eq.~(5),
  (34)]{gjerde2023directional}. We next uniquely assign the flow rate
$\langle Q'_n \rangle$ to each of the branches $\Lambda_i$ that form
the path $E_n^j$, thus yielding $\langle Q'_i \rangle$ for each
perivascular segment $\Lambda_i$ (except those ignored above, for
which we set a flow rate of zero). Finally, we define the mean
longitudinal perivascular velocity induced by the peristaltic wave
$\langle \hat{u}^x_i \rangle$ by dividing the flow rate $\langle Q'_i
\rangle$ by the cross-section area $A_i = \pi ({R_2^i}^2 - {R_1^i}^2)
= \pi (\beta_i^2 - 1) (R_1^i)^2$.

>>>>

@@

This theoretical formalism~\cite{gjerde2023directional} is defined
relative to a network in the form of a directed, bifurcating tree with
a single supply node/root $i_0$. To extend to a network of cerebral
arteries with multiple supply nodes (such as the basilar and two
internal carotid arteries in the current data
set~\cite{hodneland2019new}), we separate the network $\Lambda$ into
edge-disjoint subnetworks $\Lambda^1, \Lambda^2, \Lambda^3$, one for
each of the supply nodes (\Cref{fig:results1}A, C,
\Cref{fig:app:peristalsis}). Each node is assigned to the subnetwork
associated with the nearest supply node, and edges between nodes are
preserved. Next, we compute a minimal, bifurcating and directed tree
representation of each subnetwork: $\mathcal{T}^1, \mathcal{T}^2,
\mathcal{T}^3$. Each tree $\mathcal{T}^j = \cup E_n^j$ consists of the
subset of the nodes from $\Lambda^j$ that have degree 1 (are leaf or
root nodes) or degree $3$ (are true bifurcation points), and each path
between nodes with degree $\not = 2$ in $\Lambda^j$ is represented in
$\mathcal{T}^j$ by an edge $E^j$ with edge length $L$ corresponding to
the total length of the original path and edge radius $R^j$ as the
average of the path radii.

For each subtree $\mathcal{T}^j = \cup_n E_n^j$, we compute the
time-averaged downstream flow rate $\langle Q'_n \rangle$ induced by
the vascular wall motion for each edge $n$ via~\cite[eq.~(5),
  (34)]{gjerde2023directional}. We next uniquely assign the flow rate
$\langle Q'_n \rangle$ to each of the branches $\Lambda_i$ that form
the path $E_n^j$, thus yielding $\langle Q'_i \rangle$ for each
perivascular segment $\Lambda_i$ (except those ignored above, for
which we set a flow rate of zero). Finally, we define the mean
longitudinal perivascular velocity induced by the peristaltic wave
$\langle \hat{u}^x_i \rangle$ by dividing the flow rate $\langle Q'_i
\rangle$ by the cross-section area $A_i = \pi ({R_2^i}^2 - {R_1^i}^2)
= \pi (\beta_i^2 - 1) {R_1^i}^2$.
